Splints of root systems for special Lie subalgebras

### 2015-02-10 AffineLieAlgebras/Splints/special-paper.tex
JPhysA submission 21.01.2015

\documentclass{article}

\usepackage{amsfonts}
\usepackage{pb-diagram}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{amsmath,amssymb,amsthm}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}

@@

\title{Splints of root systems for special Lie subalgebras} % Тут надо придумать хорошее название

\author{V.D.~Lyakhovsky$^1$, A.A.~Nazarov$^{1,2}$, P.I.~Kakin$^{1}$\\
  {\small $^1$ Department of High-energy and elementary particle physics,}\\ {\small St Petersburg State University}\\
  {\small 198904, Saint-Petersburg, Russia,}\\
  {\small $^{2}$ e-mail: antonnaz@gmail.com}}
%%  {\small$^{2}$ Chebyshev Laboratory,}
%%  {\small Department of Mathematics and Mechanics, SPb State University}\\
%%  {\small 199178, Saint-Petersburg, Russia}}
\date{}
\maketitle

\begin{abstract}
  Splint is a decomposition of root system into union of root systems. Splint of root system for
  simple Lie algebra appears naturally in studies of (regular) embeddings of reductive subalgebras.
  Splint can be used to construct branching rules. We consider special embedding of Lie subalgebra
  to Lie algebra. We classify projections of algebra root systems using extended Dynkin diagrams and
  single out the conditions of splint appearance and coincidence of branching coefficients with
  weight multiplicities. While such a coincidence is not very common it is connected with Gelfand-Tsetlin basis. 
\end{abstract}

@@

The notion of splint was introduced by David Richter in the paper \cite{richter2008splints}. Splint
is the decomposition of root system into disjoint union of images of two or more embeddings of some
other root systems. Embedding $\phi$ of a root system $\Delta_1$ into a root system $\Delta$ is a
bijective map of roots of $\Delta_{1}$ to a (proper) subset of $\Delta$ that commutes with vector
composition law in $\Delta_{1}$ and $\Delta$.
\begin{equation*}
\phi:\Delta_1 \longrightarrow \Delta
\end{equation*}
\begin{equation*}
\phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,
\,\,\, \alpha,\beta \in \Delta_1
\end{equation*}

@@

Why one would consider non angle-preserving maps of root systems? Additive properties of root system
determine the structure of Verma module:
    \begin{equation*}
      M^{\mu}=U(\gf)\underset{U(\bff_{+})}{\otimes} D^{\mu}(\bff_{+})=\{(E^{-\alpha_{1}})^{n_{1}}\dots (E^{-\alpha_{s}})^{n_{s}} \left|v_{\mu}\right>\}_{\alpha_{i}\in\Delta^{+}}^{n_{i}=0,1,\dots}
    \end{equation*}
Here $\bff_{+}$ is the Borel subalgebra of $\gf$, $D^{\mu}$ its one-dimensional representation,
$\left|v_{\mu}\right>$ is the highest weight vector and $E^{-\alpha_{j}}$ are lowering operators
that are in correspondence with positive roots $\alpha_{j}\in \Delta^{+}$. 
Weyl character formula expresses character of irreducible representation as a combination of
characters of Verma modules: 
\begin{equation*}
  \mathrm{ch} L^{\mu}=\frac{\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}}{\sum_{w\in W}\varepsilon(w) e^{w\rho-\rho}}=\sum_{w\in W} \varepsilon(w)\; \mathrm{ch} M^{w(\mu+\rho)-\rho}
\end{equation*}
Here the sum is over elements of the Weyl group and their actions $w\triangleright \mu
=w(\mu+\rho)-\rho$ depend on angles between roots. But we can rewrite Weyl character formula
representing Weyl group elements as the products of reflections $s_{\alpha}, \alpha \in S$ in
hyperplanes orthogonal to simple roots: $w=s_{\alpha_{1}}\cdot s_{\alpha_{2}}\dots$. Reflections act
on root system by permutations, so the composition of such an action with embedding $\phi$ is easily
obtained. The orbit of Weyl groups action $w\triangleright \mu, w\in W$, can be constructed by
subtractions of roots from the highest weight $\mu$. Let's denote Dynkin labels of $\mu$ by
$(\mu_{1},\dots \mu_{r})$. Then $\mu-\mu_{1}\alpha_{1}, \mu-\mu_{2}\alpha_{2},\dots,
\mu-\mu_{r}\alpha_{r}, \alpha_{i}\in S$, are the points of the orbit that can be obtained by
elementary reflections $s_{\alpha_{i}}, \alpha_{i}\in S$. Next set of points that are obtained by
two consecutive reflections $s_{\alpha_{i}}\cdot s_{\alpha_{j}}\triangleright\mu$ are obtained as
the subtraction $\mu-\mu_{i}\alpha_{i}-\mu_{j} (s_{\alpha_{i}}\alpha_{j})$. Continuing this way we
construct the image of singular element $\Psi=\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}$
after the embedding $\phi$ \cite{2011arXiv1111.6787L}. The multiplicities in the character of Verma
module are unchanged by the embedding since they are determined by additive properties of the roots.
So we see that weight multiplicities of irreducible modules are also preserved by the embedding.

Root system of regular subalgebra is contained in the root system of algebra so splint is useful in
computation of branching coefficients.

In the paper \cite{2011arXiv1111.6787L} it was proven that the existence of splint leads to the
coincidence of branching coefficients with weight multiplicities under certain conditions. We denote
Lie algebra by $\gf$ and consider it's subalgebra $\af$. If $\af$ is a regular subalgebra, its root
system $\Delta_{\af}$ is contained in $\Delta_{\gf}$. Branching coefficients $b^{(\mu)}_{\nu}$
appear in decomposition of irreducible representation $L^{\mu}_{\gf}$ of $\gf$ to the sum of
irreducible representations of $\af$:
\begin{equation}
  \label{eq:1}
  L^{\mu}_{\gf}=\bigoplus_{\nu} b^{(\mu)}_{\nu} L^{\nu}_{\af}
\end{equation}

Assume that root system $\Delta_{\gf}$ splinters as $\Delta_{\gf}=\Delta_{\af} \cup
\phi(\Delta_{\sfr})$, where $\phi$ is an embedding of root system $\Delta_{\sfr}$ of some semisimple
Lie algebra $\sfr$. Then branching coefficients $b^{(\mu)}_{\nu}$ for the reduction
$L^{(\mu)}_{\gf\downarrow \af}$ coincide with weight multiplicities $m^{(\tilde \mu)}_{\nu}$ in
$\sfr$-representations provided certain technical condition holds \cite{2011arXiv1111.6787L}.
Highest weight $\tilde\mu$ of $\sfr$-representation is calculated from Dynkin labels of highest
weight $\mu$ of $\gf$-module.

In the next section \ref{sec:splints-root-systems} we review the classification of splints for simple
root systems and results for branching coefficients. Then we move to the study of special embeddings
which is the main subject of the present paper. We consider special embeddings of Lie subalgebras
into a Lie algebra. In this case root system of the subalgebra is not contained in the root system of
the algebra. So the original motivation for splints is not applicable. But we can consider the
projection of root system of the algebra on the root space of the subalgebra. Such a projection is not a
root system anymore, but it satisfies milder conditions (Section \ref{sec:spec-embedd-proj}). It is
possible to classify most projections using Dynkin diagrams augmented with multiplicities. We then
define splint for such 'weak' root systems and state the conditions of its appearance and
implications for the calculation of branching coefficients (Section \ref{sec:splints-spec-embedd}).

Use of representation theory of the subalgebra allows us to classify all the splints for the projections
of algebra root system. We also apply this method to metric splints and regular subalgebras and get
a unified treatment. We obtain Gelfand-Tsetlin rules for regular and special embeddings this way.

@@

\section{Splints of root systems and regular subalgebras}
\label{sec:splints-root-systems}
The notion and classification of
splints for simple root systems were introduced in the paper \cite{richter2008splints}.

%% Embedding $\phi$ of a root system $\Delta_1$ into a root system
%% $\Delta$ is a bijective map of roots of $\Delta_{1}$ to a (proper)
%% subset of $\Delta$ that commutes with vector composition law in
%% $\Delta_{1}$ and $\Delta$.
%% \begin{equation*}
%% \phi:\Delta_1 \longrightarrow \Delta
%% \end{equation*}
%% \begin{equation*}
%% \phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,
%% \,\,\, \alpha,\beta \in \Delta_1
%% \end{equation*}
%% 
%% Note that the image $Im(\phi)$ must not inherit the root system
%% properties except the addition rules equivalent to the addition
%% rules in $\Delta_{1}$ (for pre-images). Two embeddings $\phi_1$ and $\phi_2$  
%% can splinter $\Delta$  when the latter can be presented 
%% as a disjoint union of images $Im(\phi_1)$ and $Im(\phi_2)$.   
%% 
%% The classification is shown in table
%% 
Consider a simple Lie algebra $\mathfrak{g}$ and its regular subalgebra $%
\mathfrak{a}\hookrightarrow \mathfrak{g}$ such that $\mathfrak{a}$
is a
reductive subalgebra $\mathfrak{a \subset g}$ with correlated root spaces: $%
\mathfrak{h}_{\mathfrak{a}}^{\ast }\subset \mathfrak{h}_{\mathfrak{g }%
}^{\ast }$. Let $\mathfrak{a}^{\mathfrak{s}}$ be a semisimple summand of
$\mathfrak{a}$,
this means that $\mathfrak{a}=\mathfrak{a}^{\mathfrak{s}} \oplus \mathfrak{u}(1)\oplus %
\mathfrak{u}(1)\oplus \dots$. We shall consider $\mathfrak{a}^{\mathfrak{s}}$
to be a proper regular subalgebra and $\mathfrak{a}$ to be the
maximal subalgebra with $\mathfrak{a}^{\mathfrak{s}}$ fixed that is the rank
$r$ of $\frak{a}$ is equal to that of $\mathfrak{g}$.

Computation of branching coefficients relies on roots $\Delta_{\gf}\setminus \Delta_{\af}$
\cite{2010arXiv1007.0318L} so splint  is naturally connected to branching coefficients for regular
subalgebra.

Only three types of splints are
injective and thus are naturally connected to branching \cite{richter2008splints}:
\begin{equation}
\label{eq:1}
\begin{array}{cc||c|c}
\hbox{type} & \hspace{0.25in}\Delta \hspace{0.25in} & \hspace{0.25in}\Delta
_{\frak{a}}\hspace{0.25in} & \hspace{0.25in}\Delta _{\sfr}\hspace{0.25in}
\\ \hline\hline
\hbox{(i)} & G_{2} & A_{2} & A_{2} \\
& F_{4} & D_{4} & D_{4} \\ \hline
\hbox{(ii)} & B_{r}(r\geq 2) & D_{r} & \oplus ^{r}A_{1} \\
(*)& C_{r}(r\geq 3) & \oplus ^{r}A_{1} &  D_{r} \\ \hline
\hbox{(iii)} & A_{r}(r\geq 2) & A_{r-1}\oplus u\left( 1\right)  & \oplus
^{r}A_{1} \\
& B_{2} & A_{1}\oplus u\left( 1\right)  & A_{2}
\end{array}
\end{equation}

@@

Note that in the case of $C_{r}$-series (marked with the star in table \eqref{eq:1}) metrical
embedding $\Delta_{D_{r}}\to \Delta_{C_{r}}$ does not lead to the appearance of regular subalgebra
$\mathfrak{a}$ (See \cite{dynkin1952semisimple}).

@@

\begin{Prop}
There is a decomposition of a singular element of the algebra $\gf$ into singular elements of the subalgebra $\sfr$:
\begin{equation}
 \Phi ^{\mu }_{\gf} = \sum_{w\in W_{\frak{a}}}\epsilon\left( w\right)w\left( e^{\mu +\rho _{\gf}}\phi\left( e^{-\widetilde{\mu }}\Psi^{\widetilde{\mu }}_{\sfr}\right)\right),
\end{equation}
where $\epsilon\left( w\right)$ is a determinant of the element $w$ of the Weyl group $W_{\af}$ of
the subalgebra $\af$. The action of Weyl group on the weights is extended to the algebra of formal
exponents by the rule $w\left(e^{\nu}\right)=e^{w\nu}, w\in W$. Hence, the multiplicity $M_{\left(
      \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}$ of the weight $\widetilde{\nu }$ from the
  weight diagram of the algebra $\sfr$ with the highest weight $\tilde\mu$ defines the branching
  coefficient $b_{\nu }^{(\mu )}$ for the highest weight $\nu =\left( \mu -\phi \left(
      \widetilde{\mu }-\widetilde{\nu }\right) \right) $:
\begin{equation}
b_{\left( \mu -\phi \left( \widetilde{\mu }-\widetilde{\nu }\right) \right)
}^{(\mu )}=M_{\left( \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}. 
\label{bran1}
\end{equation}
\end{Prop} 
%% Rewrite this property informally

@@

The study of specialy embedded subalgebras traces back to fundamental papers by Eugene Dynkin
\cite{dynkin1952semisimple,dynkin1952maximal} where he called such subalgebras ``S-subalgebras'' to
distinguish from regular or ``R-subalgebras'' that have root systems obtained by dropping some roots
of the algebra root system. Thus regular subalgebras are easy to construct by dropping nodes from
extended Dynkin diagram of the algebra, while the case of special subalgebras is more difficult. Complete
classification for exceptional Lie algebras was obtained recently in \cite{minchenko2006semisimple}.
The algorithm of the special subalgebras construction is available in GAP package but still requires
manual intervention \cite{de2011constructing}.

Assume that Lie algebra $\gf$ is simple. Let's denote a subalgebra by
$\af$. We denote corresponding Cartan subalgebras by $\hfg$ and $\hfa$ and identify them with the dual
spaces $\hfg^{*}$, $\hfa^{*}$ using Killing forms of $\gf$ and $\af$.

To construct an embedding $\af\to\gf$ consider some representation $L^{\nu}_{\af}$ as a subspace of
Lie algebra $\gf$. Then one needs to check that the generators of $\af$ can be presented as
linear combinations of generators of $\gf$. We can identify Cartan subalgebra $\hfa\subset \af$ with
dual space $\hfa^{*}$ using Killing form. Then the root system $\Delta_{\gf}$ of $\gf$ can be
projected to $\hfa^{*}$ using the expression of $\hfa$-generators through $\hfg$-generators.

@@

\begin{theorem}\label{dyn1}
  Every special subalgebra $\af$ of a semisimple Lie algebra $\gf$ is integer, i.e. the projections of
  roots of $\gf$ to $\hfa^{*}$ are linear combinations of simple roots of $\af$ with integer
  coefficients \cite{dynkin1952semisimple}
\end{theorem}

@@

\begin{figure}[h!bt]
  \noindent\centering{
   \includegraphics[width=80mm]{Special-A2-D4-all-2}
  }
  \caption{Projection of $D_{4}$ ($so(8)$)-root system onto the root space of the special subalgebra $A_{2}$
    ($su(3)$). Note that this roots system is $G_{2}$ but with non-trivial multiplicities. }
 \label{fig:d4-a2_splint}
\end{figure}

@@

We want to classify all the cases when branching coefficients coincide with weight multiplicities of
some other algebra representations. Such a coincidence is possible when the projection $\Delta'$ of
$\gf$ root system admits a splint $\Delta'=\varphi_{\af}(\Delta_{\af})\cup
\varphi_{\sfr}(\Delta_{\sfr})$, where $\varphi_{\af}$ and $\varphi_{\sfr}$ are the embeddings of
corresponding root systems. The embedding $\varphi_{\af}$ is metric and trivial. Moreover, the
projection of singular element $\pia\left(\Psi^{\mu}_{\gf}\right)$ should admit a decomposition into
linear combination of singular elements of irreducible representations of $\sfr$:
\begin{equation}
  \label{eq:4}
  \pia\left(\Psi^{\mu}_{\gf}\right)=\sum_{\nu} \varkappa_{\nu}e^{\nu}\Psi^{\tilde\mu}_{\sfr},
\end{equation}
with integer coefficients $\varkappa_{\nu}$, where the sum is over the set of weights $\nu$ that
will be determined later. First we classify all the splints of the projection of a root system into
a union of simple root systems for special embeddings, and then discuss the decomposition of singular
elements.

Projection of the $\gf$ root system $\Delta'$ coincides with the projection of the weight diagram of the
adjoint representation of algebra $\gf$ with the exception of zero weight. The adjoint representation of
$\gf$ contains adjoint representation of $\af$ and can be decomposed as
\begin{equation}
  \label{eq:3}
  \mathrm{ad}_{\gf}=\mathrm{ad}_{\af}\oplus M_{\af}^{\chi},
\end{equation}
where $M^{\chi}_{\af}$ is not necessary irreducible representation of $\af$ called
``characteristic'' in \cite{dynkin1952semisimple}.

The weight diagram of $M^{\chi}_{\af}$ coincides with the root system $\Delta_{\sfr}$ with the exception of
zero weight. So we need to find all the representations for all simple Lie algebras $\af$, such that
their weight diagrams are root systems after the exclusion of zero weight. In order to do so we must
note, that all the weights of $M^{\chi}_{\af}$ should have length not less than length of shortest
root of $\af$, since $\af$ is an integer subalgebra of $\gf$ \ref{dyn1}. Moreover, $M^{\chi}_{\af}$
should contain weights of at most two different lengths.

If the root system $\Delta_{\gf}$ contains roots $\alpha: \alpha\perp \beta,\; \forall \beta\in
\Delta_{\af},$ orthogonal to the root system of the subalgebra, one cannot proceed with the decomposition
of singular element \eqref{eq:4}. The elements $e^{\nu}$ of $\pia\left(\Psi^{\mu}_{\gf}\right)$
should be augmented with the dimensions of representations of algebra $\afb$ spanned by generators
corresponding to orthogonal roots \cite{2010arXiv1007.0318L}. Then there will be non-trivial
multiplicities in the right hand side of \eqref{eq:4} and no coincidence of branching coefficients
with weight multiplicities of $\sfr$-representations. One can write more cumbersome relation between
multiplicities and branching coefficients, but it is out of scope for the present paper.

>>>>

\documentclass[12pt]{iopart}

%\usepackage{amsfonts}
\usepackage{pb-diagram}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{iopams,setstack}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}

@@

\ftc{Splints of root systems for special Lie subalgebras}

\author{V.D.~Lyakhovsky$^1$, A.A.~Nazarov$^{1,2}$, P.I.~Kakin$^{1}$}
\address{ $^1$ Department of High-energy and elementary particle physics,  St Petersburg State University, 198904, Saint-Petersburg, Russia}
\ead{ $^{2}$ e-mail: antonnaz@gmail.com}

\begin{abstract}
  Splint is a decomposition of root system into a union of root systems. Splint of root system for a
  simple Lie algebra appears naturally in the studies of (regular) embeddings of reductive
  subalgebras. Splint can be used to construct branching rules. We consider a special embedding of a
  Lie subalgebra into a simple Lie algebra. We classify the projections of algebra root systems and
  single out the conditions of splint appearance and coincidence of branching coefficients with
  weight multiplicities. While such a coincidence is not very common it is connected with
  Gelfand-Tsetlin basis.

\noindent{\it Keywords\/}: Lie algebra, special subalgebra, branching, weight multiplicity, splint
\end{abstract}

\submitto{\jpa}

@@

The notion of splint was introduced by David Richter in the paper \cite{richter2008splints}. Splint
is a decomposition of a root system into a disjoint union of the images of two or more embeddings of some
other root systems. An embedding $\phi$ of a root system $\Delta_1$ into a root system $\Delta$ is a
bijective map of roots of $\Delta_{1}$ to a (proper) subset of $\Delta$ that commutes with the vector
composition law in $\Delta_{1}$ and $\Delta$.
\begin{equation*}
\phi:\Delta_1 \longrightarrow \Delta
\end{equation*}
\begin{equation*}
\phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,
\,\,\, \alpha,\beta \in \Delta_1
\end{equation*}

@@

Why one would consider non angle-preserving maps of root systems? Additive properties of the root system
determine the structure of the Verma module:
    \begin{equation*}
      M^{\mu}=U(\gf)\underset{U(\bff_{+})}{\otimes} D^{\mu}(\bff_{+})=\{(E^{-\alpha_{1}})^{n_{1}}\dots (E^{-\alpha_{s}})^{n_{s}} \left|v_{\mu}\right>\}_{\alpha_{i}\in\Delta^{+}}^{n_{i}=0,1,\dots}
    \end{equation*}
Here $\bff_{+}$ is the Borel subalgebra of $\gf$, $D^{\mu}$ is its one-dimensional representation,
$\left|v_{\mu}\right>$ is the highest weight vector and $E^{-\alpha_{j}}$ are lowering operators
that are in correspondence with positive roots $\alpha_{j}\in \Delta^{+}$. 
The Weyl character formula expresses a character of an irreducible representation as a combination of
the characters of the Verma modules: 
\begin{equation*}
  \mathrm{ch} L^{\mu}=\frac{\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}}{\sum_{w\in W}\varepsilon(w) e^{w\rho-\rho}}=\sum_{w\in W} \varepsilon(w)\; \mathrm{ch} M^{w(\mu+\rho)-\rho}
\end{equation*}
Here the sum is over elements of the Weyl group and their actions $w\triangleright \mu
=w(\mu+\rho)-\rho$ depend on angles between roots. But we can rewrite the Weyl character formula
representing the Weyl group elements as the products of the reflections $s_{\alpha}, \alpha \in S$ in
the hyperplanes orthogonal to the simple roots: $w=s_{\alpha_{1}}\cdot s_{\alpha_{2}}\dots$. The reflections act
on the root system by permutations, so the composition of such an action with the embedding $\phi$ is easily
obtained. The orbit of the Weyl group action $w\triangleright \mu, w\in W$, can be constructed by
subtractions of roots from the highest weight $\mu$. Let's denote the Dynkin labels of $\mu$ by
$(\mu_{1},\dots \mu_{r})$. Then $\mu-\mu_{1}\alpha_{1}, \mu-\mu_{2}\alpha_{2},\dots,
\mu-\mu_{r}\alpha_{r}, \alpha_{i}\in S$, are the points of the orbit that can be obtained by
the elementary reflections $s_{\alpha_{i}}, \alpha_{i}\in S$. The next set of points that are obtained by
two consecutive reflections $s_{\alpha_{i}}\cdot s_{\alpha_{j}}\triangleright\mu$ are obtained as
the subtraction $\mu-\mu_{i}\alpha_{i}-\mu_{j} (s_{\alpha_{i}}\alpha_{j})$. Continuing this way we
construct the image of the singular element $\Psi=\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}$
after the embedding $\phi$ \cite{2011arXiv1111.6787L}. The multiplicities in the character of the Verma
module are unchanged by the embedding since they are determined by the additive properties of the roots.
So we see that the weight multiplicities of irreducible modules are also preserved by the embedding.

The root system of the regular subalgebra is contained in the root system of the algebra so the splint is useful in the
computation of branching coefficients.

In the paper \cite{2011arXiv1111.6787L} it was proven that the existence of splint leads to the
coincidence of branching coefficients with weight multiplicities under certain conditions. We denote a
Lie algebra by $\gf$ and consider it's subalgebra $\af$. If $\af$ is a regular subalgebra, its root
system $\Delta_{\af}$ is contained in $\Delta_{\gf}$. Branching coefficients $b^{(\mu)}_{\nu}$
appear in the decomposition of an irreducible representation $L^{\mu}_{\gf}$ of $\gf$ to the sum of
irreducible representations of $\af$:
\begin{equation}
  \label{eq:1}
  L^{\mu}_{\gf}=\bigoplus_{\nu} b^{(\mu)}_{\nu} L^{\nu}_{\af}
\end{equation}

Assume that the root system $\Delta_{\gf}$ splinters as $\Delta_{\gf}=\Delta_{\af} \cup
\phi(\Delta_{\sfr})$, where $\phi$ is an embedding of a root system $\Delta_{\sfr}$ of some semisimple
Lie algebra $\sfr$. Then branching coefficients $b^{(\mu)}_{\nu}$ for the reduction
$L^{(\mu)}_{\gf\downarrow \af}$ coincide with the weight multiplicities $m^{(\tilde \mu)}_{\nu}$ in the $\sfr$-representations provided certain technical condition holds \cite{2011arXiv1111.6787L}.
The highest weight $\tilde\mu$ of the $\sfr$-representation is calculated from the Dynkin labels of the highest
weight $\mu$ of the $\gf$-module.

In the next section \ref{sec:splints-root-systems} we review the classification of splints for simple
root systems and the results for branching coefficients. Then we move to the study of special embeddings
which is the main subject of the present paper. We consider special embeddings of Lie subalgebras
into a Lie algebra. In this case the root system of the subalgebra is not contained in the root system of
the algebra. So the original motivation for splints is not applicable. But we can consider the
projection of the root system of the algebra on the root space of the subalgebra. Such a projection is not a
root system anymore, but it satisfies milder conditions (Section \ref{sec:spec-embedd-proj}). It is
possible to classify the splints of the projection using the fact that the adjoint representation of the algebra is decomposed into the direct sum of representations of the subalgebra. We show that if there is a splint then branching coefficients coincide with weight multiplicities of representations of another Lie algebra (Section \ref{sec:splints-spec-embedd}). Thus in this case branching coefficients are very easy to compute.

The use of representation theory of the subalgebra allows us to classify all the splints for the projections
of the algebra root system. We also apply this method to metric splints and regular subalgebras and get
a unified treatment. We obtain the Gelfand-Tsetlin rules for regular and special embeddings this way.

@@

\section{Splints of root systems and regular subalgebras}
\label{sec:splints-root-systems}
The notion and the classification of
splints for simple root systems were introduced in the paper \cite{richter2008splints}.

Consider a simple Lie algebra $\mathfrak{g}$ and its regular subalgebra $%
\mathfrak{a}\hookrightarrow \mathfrak{g}$ such that $\mathfrak{a}$
is a
reductive subalgebra $\mathfrak{a \subset g}$ with correlated root spaces: $%
\mathfrak{h}_{\mathfrak{a}}^{\ast }\subset \mathfrak{h}_{\mathfrak{g }%
}^{\ast }$. Let $\mathfrak{a}^{\mathfrak{s}}$ be a semisimple summand of
$\mathfrak{a}$,
this means that $\mathfrak{a}=\mathfrak{a}^{\mathfrak{s}} \oplus \mathfrak{u}(1)\oplus %
\mathfrak{u}(1)\oplus \dots$. We shall consider $\mathfrak{a}^{\mathfrak{s}}$
to be a proper regular subalgebra and $\mathfrak{a}$ to be the
maximal subalgebra with $\mathfrak{a}^{\mathfrak{s}}$ fixed, so that the rank
$r$ of $\af$ is equal to that of $\mathfrak{g}$.

The computation of branching coefficients relies on the roots $\Delta_{\gf}\setminus \Delta_{\af}$
\cite{2010arXiv1007.0318L} so splint  is naturally connected to branching coefficients for regular
subalgebras.

Only three types of splints are
injective and thus are naturally connected to branching \cite{richter2008splints}:
\begin{equation}
\label{tab:1}
\begin{array}{cc||c|c}
\hbox{type} & \hspace{0.25in}\Delta \hspace{0.25in} & \hspace{0.25in}\Delta
_{\frak{a}}\hspace{0.25in} & \hspace{0.25in}\Delta _{\sfr}\hspace{0.25in}
\\ \hline\hline
\hbox{(i)} & G_{2} & A_{2} & A_{2} \\
& F_{4} & D_{4} & D_{4} \\ \hline
\hbox{(ii)} & B_{r}(r\geq 2) & D_{r} & \oplus ^{r}A_{1} \\
(*)& C_{r}(r\geq 3) & \oplus ^{r}A_{1} &  D_{r} \\ \hline
\hbox{(iii)} & A_{r}(r\geq 2) & A_{r-1}\oplus u\left( 1\right)  & \oplus
^{r}A_{1} \\
& B_{2} & A_{1}\oplus u\left( 1\right)  & A_{2}
\end{array}
\end{equation}

@@

Note that in the case of $C_{r}$-series (marked with the star in table \ref{tab:1}) the metrical
embedding $\Delta_{D_{r}}\to \Delta_{C_{r}}$ does not lead to the appearance of the regular subalgebra
$\mathfrak{a}$ (See \cite{dynkin1952semisimple}).

@@

\begin{Prop}
There is a decomposition of a singular element of the algebra $\gf$ into singular elements of the subalgebra $\sfr$:
\begin{equation}
 \Phi ^{\mu }_{\gf} = \sum_{w\in W_{\frak{a}}}\epsilon\left( w\right)w\left( e^{\mu +\rho _{\gf}}\phi\left( e^{-\widetilde{\mu }}\Psi^{\widetilde{\mu }}_{\sfr}\right)\right),
\end{equation}
where $\epsilon\left( w\right)$ is a determinant of the element $w$ of the Weyl group $W_{\af}$ of
the subalgebra $\af$. The action of the Weyl group on the weights is extended to the algebra of formal
exponents by the rule $w\left(e^{\nu}\right)=e^{w\nu}, w\in W$. Hence, the multiplicity $M_{\left(
      \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}$ of the weight $\widetilde{\nu }$ from the
  weight diagram of the algebra $\sfr$ with the highest weight $\tilde\mu$ defines the branching
  coefficient $b_{\nu }^{(\mu )}$ for the highest weight $\nu =\left( \mu -\phi \left(
      \widetilde{\mu }-\widetilde{\nu }\right) \right) $:
\begin{equation}
b_{\left( \mu -\phi \left( \widetilde{\mu }-\widetilde{\nu }\right) \right)
}^{(\mu )}=M_{\left( \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}. 
\label{bran1}
\end{equation}
\end{Prop} 
%% Rewrite this property informally

@@

The study of specialy embedded subalgebras traces back to the fundamental papers by Eugene Dynkin
\cite{dynkin1952semisimple,dynkin1952maximal} where he called such subalgebras ``S-subalgebras'' to
distinguish from regular or ``R-subalgebras'' that have root systems obtained by dropping some roots
of the algebra root system. Thus regular subalgebras are easy to construct by dropping nodes from
the extended Dynkin diagram of the algebra, while the case of special subalgebras is more difficult. The complete
classification for exceptional Lie algebras was obtained recently in \cite{minchenko2006semisimple}.
The algorithm of the special subalgebras construction is available in GAP package but still requires
a manual intervention \cite{de2011constructing}.

Assume that the Lie algebra $\gf$ is simple. Let's denote a subalgebra by
$\af$. We denote corresponding Cartan subalgebras by $\hfg$ and $\hfa$ and identify them with the dual
spaces $\hfg^{*}$, $\hfa^{*}$ using the Killing forms of $\gf$ and $\af$.

To construct an embedding $\af\to\gf$ consider some representation $L^{\nu}_{\af}$ as a subspace of
Lie algebra $\gf$. Then one needs to check that the generators of $\af$ can be presented as
linear combinations of generators of $\gf$. We can identify the Cartan subalgebra $\hfa\subset \af$ with
the dual space $\hfa^{*}$ using the Killing form. Then the root system $\Delta_{\gf}$ of $\gf$ can be
projected to $\hfa^{*}$ using the expression of $\hfa$-generators through $\hfg$-generators.

@@

\begin{theorem}\label{dyn1}
  Every special subalgebra $\af$ of a semisimple Lie algebra $\gf$ is integer, i.e. the projections of
  the roots of $\gf$ to $\hfa^{*}$ are linear combinations of the simple roots of $\af$ with integer
  coefficients \cite{dynkin1952semisimple}
\end{theorem}

@@

\begin{figure}[h!bt]
  \noindent\centering{
   \includegraphics[width=40mm]{Special-A2-D4-all-2}
  }
  \caption{Projection of $D_{4}$ ($so(8)$)-root system onto the root space of the special subalgebra $A_{2}$
    ($su(3)$). Note that this root system is $G_{2}$ but with non-trivial multiplicities. }
 \label{fig:d4-a2_splint}
\end{figure}

@@

We want to classify all the cases when branching coefficients coincide with weight multiplicities of
some other algebra representations. Such a coincidence is possible when the projection $\Delta'$ of
the $\gf$ root system admits a splint $\Delta'=\varphi_{\af}(\Delta_{\af})\cup
\varphi_{\sfr}(\Delta_{\sfr})$, where $\varphi_{\af}$ and $\varphi_{\sfr}$ are the embeddings of
the corresponding root systems. The embedding $\varphi_{\af}$ is metric and trivial. Moreover, the
projection of the singular element $\pia\left(\Psi^{\mu}_{\gf}\right)$ should admit a decomposition into
a linear combination of singular elements of irreducible representations of $\sfr$:
\begin{equation}
  \label{eq:4}
  \pia\left(\Psi^{\mu}_{\gf}\right)=\sum_{\nu} \varkappa_{\nu}e^{\nu}\Psi^{\tilde\mu}_{\sfr},
\end{equation}
with integer coefficients $\varkappa_{\nu}$, where the sum is over the set of weights $\nu$ that
will be determined later. Such a decomposition of the singular element is very similar to the branching of Weyl group orbits considered in the papers \cite{larouche2011branching,larouche2009branching}. First we classify all the splints of the projection of a root system into
a union of simple root systems for special embeddings, and then discuss the decomposition of singular
elements.

Projection of the $\gf$ root system $\Delta'$ coincides with the projection of the weight diagram of the
adjoint representation of the algebra $\gf$ with the exception of zero weight. The adjoint representation of
$\gf$ contains the adjoint representation of $\af$ and can be decomposed as
\begin{equation}
  \label{eq:3}
  \mathrm{ad}_{\gf}=\mathrm{ad}_{\af}\oplus M_{\af}^{\chi},
\end{equation}
where $M^{\chi}_{\af}$ is not necessary irreducible representation of $\af$ called
``characteristic'' in \cite{dynkin1952semisimple}.

The weight diagram of $M^{\chi}_{\af}$ coincides with the root system $\Delta_{\sfr}$ with the exception of
zero weight. So we need to find all the representations for all simple Lie algebras $\af$, such that
their weight diagrams are root systems after the exclusion of zero weight. In order to do so we must
note, that all the weights of $M^{\chi}_{\af}$ should have a length not less than the length of the shortest
root of $\af$, since $\af$ is an integer subalgebra of $\gf$ \ref{dyn1}. Moreover, $M^{\chi}_{\af}$
should contain weights of at most two different lengths.
